---
title: "data-cleaning-2"
output: html_document
date: "2023-10-10"
---

# IMPORTING PACKAGES
```{r}
library(dplyr)
library(arrow)
library(tidyverse)
library(sparklyr)
library(dbplot)
```

# EDA

## Connect to spark and load accidents data
```{r}
# CONNECT
sc <- spark_connect(master = "local", version = "3.4.0")

accidents <- spark_read_parquet(sc, path = "data/us_accidents")

glimpse(accidents)
head(accidents)
```

## Distribution of Severity
```{r}
grouped_severity <- accidents |>
  group_by(Severity) |>
  summarise(Count = n()) |>
  collect()

# Visualize with Spark
grouped_severity |> 
  mutate(Severity = as.factor(Severity), # Specify Severity to be factor (categorical)  
         Severity = reorder(Severity, desc(Count))) |> # Reorder in descending Count
  ggplot(aes(x = Severity, y = Count)) +
  geom_col(fill = "deepskyblue4", width = 0.4) +
  labs(title = "Distribution of Accidents by Severity")
```
Majority of the accidents are of severity 2. If we are classifying Severity we might need to perform some resampling.

## Number of Accidents by Time of Day
```{r}
grouped_tod <- accidents |>
  group_by(TimeOfDay) |>
  summarise(Count = n()) |>
  collect()

# Visualize with Spark
grouped_tod |> 
  mutate(TimeOfDay = as.factor(TimeOfDay), # Specify TimeOfDay to be factor (categorical)  
         TimeOfDay = reorder(TimeOfDay, desc(Count))) |>  # Reorder in descending Count
  ggplot(aes(x = TimeOfDay, y = Count)) +
  geom_col(fill = "deepskyblue4", width = 0.4) + 
  labs(title = "Distribution of Accidents by Time of Day")
```
Surprisingly there are more accidents in the day.

## Top 5 Weather Conditions
```{r}
top_5_weather <- accidents |>
  group_by(Weather_Condition) |>
  summarise(Count = n()) |>
  slice_max(order_by = Count, n = 5, with_ties = FALSE) |>
  collect()

# Visualize with Spark
top_5_weather |> 
  mutate(Weather_Condition = as.factor(Weather_Condition),
         Weather_Condition = reorder(Weather_Condition, desc(Count))) |>
  ggplot(aes(x = Weather_Condition, y = Count)) +
  geom_col(fill = "deepskyblue4", width = 0.4) + 
  labs(title = "Top 5 Weather Conditions")
```

## Explore the distribution of Visibilitymi
```{r}
grouped_vis <- accidents |>
  group_by(Visibilitymi) |>
  summarise(Count = n()) |>
  collect()

# Visualize with Spark
grouped_vis |> 
  mutate(Visibilitymi = as.factor(Visibilitymi), # Specify Visibilitymi to be factor (categorical)  
         Visibilitymi = reorder(Visibilitymi, desc(Count))) |>  # Reorder in descending Count
  ggplot(aes(x = Visibilitymi, y = Count)) +
  geom_col(fill = "deepskyblue4", width = 0.4) + 
  labs(title = "Distribution of Accidents by Visibilitymi")
```

## Exploring the more severe accidents
```{r}
severe_accidents <- accidents |>
  filter(Severity > 2)

# Time of day for severe accidents
grouped_tod <- severe_accidents |>
  group_by(TimeOfDay) |>
  summarise(Count = n()) |>
  collect()

grouped_tod |> 
  mutate(TimeOfDay = as.factor(TimeOfDay), # Specify TimeOfDay to be factor (categorical)  
         TimeOfDay = reorder(TimeOfDay, desc(Count))) |>  # Reorder in descending Count
  ggplot(aes(x = TimeOfDay, y = Count)) +
  geom_col(fill = "deepskyblue4", width = 0.4) + 
  labs(title = "Distribution of Accidents by Time of Day")

# Weather conditions for severe accidents
top_5_weather <- severe_accidents |>
  group_by(Weather_Condition) |>
  summarise(Count = n()) |>
  slice_max(order_by = Count, n = 5, with_ties = FALSE) |>
  collect()

top_5_weather |> 
  mutate(Weather_Condition = as.factor(Weather_Condition),
         Weather_Condition = reorder(Weather_Condition, desc(Count))) |>
  ggplot(aes(x = Weather_Condition, y = Count)) +
  geom_col(fill = "deepskyblue4", width = 0.4) + 
  labs(title = "Top 5 Weather Conditions for Severe Accidents")

# Distribution of Visibilitymi
grouped_vis <- severe_accidents |>
  group_by(Visibilitymi) |>
  summarise(Count = n()) |>
  collect()

# Visualize with Spark
grouped_vis |> 
  mutate(Visibilitymi = as.factor(Visibilitymi), # Specify Visibilitymi to be factor (categorical)  
         Visibilitymi = reorder(Visibilitymi, desc(Count))) |>  # Reorder in descending Count
  ggplot(aes(x = Visibilitymi, y = Count)) +
  geom_col(fill = "deepskyblue4", width = 0.4) + 
  labs(title = "Distribution of Accidents by Visibilitymi for Severe Accidents")
```
Distribution of variables are similar despite filtering for more severe accidents.

## Disconnect
```{r}
# DISCONNECT 
spark_disconnect(sc)
```

